### 2019-01-28 intro.tex
ssr

\section{Introduction} 
\label{sec:intro}

\subsection{Canalyzing and Nested Canalyzing Functions}
\label{sse:ncf_def}

The following definition of \textbf{canalyzing} Boolean functions was introduced
in \cite{Kauffman-1969}.

\begin{definition}\label{def:canalyzing}
Let $X = $ $\{x_1, x_2, \ldots, x_n\}$ be a set of $n$  Boolean variables.
A Boolean function $f(x_1, x_2, \ldots, x_n)$ over $X$ is \textbf{canalyzing}
if there is a variable $x_i \in X$ and values $a$, $b$ $\in \{0,1\}$ such that 
$f(x_1, \ldots, x_{i-1}, a, x_{i+1}, \ldots, x_n)$ $~=~ b$, regardless of the values
assigned to the variables in $X - \{x_i\}$.
\end{definition}

\medskip

\noindent
\textbf{Example 1:}~ Consider the function 
$f(x_1, x_2, x_3) ~=~ \overline{x_1} \wedge (x_2 \vee \overline{x_3})$.
This function is canalyzing since if we set to $x_1 = 1$,
$f(1, x_2, x_3) ~=~ 0$, regardless of the values assigned to $x_2$ and $x_3$. \qed

\medskip

Another class of Boolean functions, called \textbf{nested canalyzing functions} (NCFs),
was introduced later in \cite{Kauffman-etal-2003} to model the behavior of
certain biological systems.
We follow the presentation in \cite{Layne-2011} in defining 
such a Boolean function. 
(For a Boolean value $b$,~ the complement is denoted by $\overline{b}$.)

\begin{definition}\label{def:nested_canalyzing}
Let $X = $ $\{x_1, x_2, \ldots, x_n\}$ denote a set of $n$  Boolean variables.
Let $\pi$ be a permutation of $\{1, 2, \ldots, n\}$.
A Boolean function $f(x_1, x_2, \ldots, x_n)$ over $X$ is \textbf{nested canalyzing}
in the variable order $x_{\pi(1)}, x_{\pi(2)}, \ldots, x_{\pi(n)}$ with
\textbf{canalyzing values} $a_1, a_2, \ldots, a_n$ and 
\textbf{canalyzed values} $b_1, b_2, \ldots, b_n$ 
if $f$ can be expressed in the following form:
\[
f(x_1, x_2, \ldots, x_n) ~=~ 
   \begin{cases}
       \:b_1 & \mathrm{if~~} x_{\pi(1)} ~=~ a_1 \\
       \:b_2 & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~}
            x_{\pi(2)} ~=~ a_2 \\
       \:\vdots & \vdots \\
       \:b_n & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~} \ldots~
             x_{\pi(n-1)} ~\neq~ a_{n-1} \mathrm{~~and~~} x_{\pi(n)} ~=~ a_n \\
       \:\overline{b_n} & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~} \ldots~
            x_{\pi(n)} ~\neq~ a_n \\
   \end{cases}
\]
\end{definition}

\medskip
%% \clearpage
For convenience, we will use a computational notation to represent NCFs
introduced in \cite{Stearns-etal-2018}.
For $1 \leq i \leq n$, line $i$ of this representation has the following form:

\medskip

\noindent
\hspace*{1.1in} $x_{\pi(i)}:~ a_i ~~\longrightarrow~~ b_i$

\medskip

\noindent 
Following the terminology in \cite{Stearns-etal-2018},
 $x_{\pi(i)}$ is the \textbf{canalyzing variable} that is
\textbf{tested} in line $i$, 
with $a_i$ and $b_i$ denoting respectively the canalyzing and 
canalyzed values in line $i$ as before,~ $1 \leq i \leq n$.
each such line is called a \textbf{rule}.
When none of the conditions ``$x_{\pi(i)} ~=~ a_i$" 
is satisfied, we have line $n+1$ with the ``Default" rule
for which the canalyzed value is~ $\overline{b_n}$:

\medskip

\noindent
\hspace*{1.1in} Default:~ $\overline{b_n}$

\medskip
\noindent
As in \cite{Stearns-etal-2018},, we will refer to the above specification
of an NCF as the \textbf{simplified representation} and assume
(without loss of generality) that each NCF is specified in this manner.
%% The simplified representation provides the following 
%% convenient computational view of an NCF.~
%% Lines defining an NCF are 
%% considered sequentially in a top-down manner.
%% The computation stops at the first line where the 
%% specified condition is satisfied, and the value of the function
%% is the canalyzed value on that line. 
We now present an example of an NCF using the two representations
mentioned above.

\medskip
\noindent
\textbf{Example 2:}~ Consider the function 
$f(x_1, x_2, x_3) ~=~ \overline{x_1} \wedge (x_2 \vee \overline{x_3})$.
This function is nested canalyzing using the identity permutation $\pi$ on $\{1,2,3\}$
with canalyzing values $1,1,0$ and canalyzed values $0, 1, 1$.
We first show how this function can be expressed using the syntax of
Definition~\ref{def:nested_canalyzing}.

\[
f(x_1, x_2, x_3) ~=~ 
   \begin{cases}
       \:0 & \mathrm{if~~} x_{1} ~=~ 1 \\
       \:1 & \mathrm{if~~} x_{1} ~\neq~ 1 \mathrm{~~and~~}
            x_{2} ~=~ 1 \\
       \:1 & \mathrm{if~~} x_{1} ~\neq~ 1 \mathrm{~~and~~}
            x_{2} ~\neq~ 1 \mathrm{~~and~~} x_{3} = 0 \\
       \:0 & \mathrm{if~~} x_{1} ~\neq~ 1 \mathrm{~~and~~}
            x_{2} ~\neq~ 1 \mathrm{~~and~~} x_{3} ~\neq~ 0 \\
   \end{cases}
\]

\medskip
\noindent
A simplified representation of the same function is as follows.

\bigskip

\noindent
\begin{tabular}{ll}
\hspace*{1.1in} & $x_1:~$  $1 ~\longrightarrow~ 0$ \\ [1ex]
\hspace*{1.1in} & $x_2:~$  $1 ~\longrightarrow~ 1$ \\ [1ex]
\hspace*{1.1in} & $x_3:~$  $0 ~\longrightarrow~ 1$ \\ [1ex]
\hspace*{1.1in} & Default:~ $0$ \\
\end{tabular}

\noindent
\qed

\iffalse
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent
\underline{\textsf{Generalized NCFs:}}~
For some problems, we will also consider a more general form
of NCFs defined below.

\begin{definition}\label{def:generalized ncf}
A {\bf generalized NCF} is a function represented as either a constant
or an NCF representation of a subset (not necessarily proper) 
of the function's variables.
\end{definition}

\noindent
\textbf{Example 3:}~ For $b \in \{0,1\}$, the constant function which takes 
on the value $b$ for every input is a generalized NCF since 
it can be represented by the following rule:

\smallskip

\hspace*{1.1in} Default:~ $b$

\smallskip

\noindent
The following generalized NCF specification for 
a function $f(x_1, x_2, x_3, x_4)$ indicates that the function
depends only on variables $x_1$ and $x_3$.

\medskip

\noindent
\begin{tabular}{ll}
\hspace*{1.1in} & $x_1:~$  $0 ~\longrightarrow~ 1$ \\ [1ex]
\hspace*{1.1in} & $x_3:~$  $1 ~\longrightarrow~ 0$ \\ [1ex]
\hspace*{1.1in} & Default:~ $1$ \\
\end{tabular}

\noindent
\qed

\bigskip

\noindent
\textbf{Note:} If a generalized NCF is not a constant, 
there is an assignment to its variables
such that the value of the function is 0,
and there is an assignment to its variables
such that the value of the function is 1.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fi

\subsection{Symmetric Boolean Functions}
\label{sse:symmetry}

A pair of variables of a Boolean function $f(x_1, x_2, \ldots, x_n)$ is
said to be \textbf{symmetric} if their values can be interchanged without
affecting the value of the function.
This form of interchange symmetry partitions the set of variables into a set of
\textbf{symmetry groups}, where the members of each group are pairwise symmetric.
Function $f$ is said to be
$r$-\textbf{symmetric} if it has at most $r$ symmetry groups.
In this case, the value of $f$
depends only on how many of the variables in each symmetry group have the value 1.
We say that $f$ is \textbf{properly} $r$-\textbf{symmetric} if
it is $r$-symmetric, but not $(r-1)$-symmetric.

In the literature (see, e.g., \cite{Crama-Hammer-2011,HT-2016,Toth-etal-1977}),
a 1-symmetric function $f$ is referred to simply
as a \textbf{symmetric} function.
Thus, if $f$ is a symmetric function, then for any
input $(a_1, a_2, \ldots, a_n)$ to $f$, where $a_i \in \{0,1\}$ for
$1 \leq i \leq n$, and any permutation $\pi$ of $\{1, 2, \ldots, n\}$,
$f(a_1, a_2, \ldots, a_n)$ = $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.

%\noindent

\subsection{Summary of Results}
\label{sse:contrib}

Our focus is on the relationship between NCFs and symmetric
Boolean functions.

\subsection{Related Work}
\label{sse:related}

The term \textbf{canalization} was first used by
Conrad Waddington (a British geneticist)
in 1942 \cite{Waddington-1942}
to express the ability of a species
to produce members with very similar observable features over a period
of time even when there are significant changes in the environment
during that period.
Since that time, the term ``canalization'' has been used to describe
the stability of a biological system with changes
in external conditions.
In 1969, Kauffman \cite{Kauffman-1969} introduced a Boolean network model
to explain the stability of gene regulatory networks.
Kauffman observed that the use of canalyzing Boolean
functions in the model captured many observed properties of
gene regulatory networks.
The subclass of NCFs was
introduced later by Kauffman et al. \cite{Kauffman-etal-2003} 
to facilitate a rigorous analysis of the Boolean network model
for gene regulatory networks.
Many researchers have studied mathematical properties of NCFs and
pointed out the importance of NCFs in modeling biological
phenomena (e.g., \cite{Layne-2011,
Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}).

There is extensive literature on the sensitivity of various classes
of Boolean functions
(e.g., \cite{Buhrman-etal-2002,Nisan-etal-1994,Zhang-2011}).
For a discussion on how the sensitivity of a Boolean function provides
an indication of its stability, the reader is referred to \cite{Kauffman-etal-2004,
Layne-2011,Layne-etal-2012}.
Li and Adeyeye \cite{Li-etal-2012} present lower and upper bounds on the
sensitivity of any NCF.
Li et al. \cite{Li-etal-2011,Li-etal-2013} conjectured that the average
sensitivity of any NCF is strictly less than  $4/3$.
This conjecture was proved by Klotz et al. \cite{Klotz-etal-2013} by establishing
a tight upper bound on the average sensitivity of any NCF.
Their methods rely on Fourier analysis of Boolean functions
\cite{Odonnell-2014}.
In \cite{Stearns-etal-2018}, 
we presented a characterization of NCFs with maximum average sensitivity.
This characterization allowed us to present an alternative proof 
of the upper bound in \cite{Klotz-etal-2013} and derive a closed
form expression for the number of such NCFs.