### 2019-02-01 abstract.tex
ssr

\vspace*{-0.05in}
\begin{center}
\textbf{Abstract}  %% ~~ \textcolor{red}{\textbf{(To be revised)}}
\end{center}

\smallskip

A class of Boolean functions, 
called \textbf{nested canalyzing functions} (NCFs),
has been used to model certain biological phenomena.
Researchers have studied some properties of these functions
such as sensitivity and stability.
In this note, we examine the relationships between NCFs and symmetric 
Boolean functions. 
A more general form of symmetric Boolean functions,
called $r$-symmetric functions (where $r$ is the symmetry level),
has also been studied in the literature.
We observe that, in general,
one cannot efficiently approximate the symmetry level of
a Boolean function to within any factor, unless \textbf{P} = \cnp. 
In contrast, we show that the symmetry level of an NCF $f$
can be easily computed given a standard representation of $f$.
We also show that for any NCF $f$ with $n$ variables, the notion of
strong asymmetry considered in the literature is equivalent to
the property that the symmetry level of $f$ is $n$.
We also identify all the Boolean functions that NCFs 
as well as symmetric.