### 2019-03-29 IEEE_TC_submission.d/summary.tex

\section{Summary}
\label{sec:summary}

We presented a characterization of when two variables
of an NCF are symmetric and used that characterization
to show that the symmetry level of an NCF can be easily computed.
We also showed that an $n$-variable NCF is strongly asymmetric
iff its symmetry level is $n$.
We presented several corollaries of this result, including
a closed form expression for the number of NCFs which are also
strongly asymmetric.
Thus, our results bring out several interesting relationships 
between NCFs and symmetric Boolean functions.